---
output: github_document
---

[![Travis-CI Build Status](https://travis-ci.org/fboehm/qtl2pleio.svg?branch=master)](https://travis-ci.org/fboehm/qtl2pleio)

[![Coverage Status](https://img.shields.io/codecov/c/github/fboehm/qtl2pleio/master.svg)](https://codecov.io/github/fboehm/qtl2pleio?branch=master)


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# qtl2pleio

The goal of qtl2pleio is to perform a likelihood ratio test that distinguishes between competing hypotheses of presence of two separate QTL (alternative hypothesis) and pleiotropy (null hypothesis) in QTL studies in multiparental populations, such as the Diversity Outbred mouse population. `qtl2pleio` data structures are those used in the `rqtl/qtl2` package. 

## Installation

You can install `qtl2pleio` from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("fboehm/qtl2pleio")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```
