---
title: "Using Condor for bootstrap analysis"
author: "Frederick Boehm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Condor for bootstrap analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  tidy = TRUE, 
  tidy.opts = list(width.cutoff = 60)
  )
```


## Overview

Performing a bootstrap analysis with qtl2pleio requires first sampling a (bivariate) phenotype from the inferred bivariate normal distribution, then a two-dimensional QTL scan over the defined scan region. Often, one wants to include more than 100 markers in a scan region. Each two-dimensional scan can take tens of minutes with a single core. Thus, performing a bootstrap analysis with, say, 1000 bootstrap samples is prohibitively time-consuming on most computers.
For this reason, we used a high-throughput computing cluster at the University of Wisconsin-Madison (where qtl2pleio was developed). The UW-Madison Center for High-Throughput Computing is available to all UW-Madison researchers. While we recognize that qtl2pleio users may not have access to the CHTC, we anticipate that some will have access to computing cl

## Directory structure, files & setup for using Condor



## Submitting jobs to Condor (at UW-Madison CHTC)






## Session info

```{r}
devtools::session_info()
```


