---
title: "Comparing covariance estimates between `limmbo2` and `gemma2`"
author: "Frederick Boehm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing covariance estimates between `limmbo2` and `gemma2`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background

[`gemma2`](https://github.com/fboehm/gemma2) is an R implementation of the C++ code for EM-based estimation of Vg and Ve. `limmbo2` is a R wrapper to the [`limmbo`](https://github.com/HannahVMeyer/limmbo) python module. 

## Goals

I want to see how similar the variance component estimates are for the two methods. Recall that it's the function `qtl2pleio::calc_covs` that I want to use to get the estimates based on gemma2's EM algorithm. `limmbo2` in the limmbo2 R package will provide both Vg and Ve.

## First, `gemma2` method:

```{r}
library(qtl2pleio)
pheno <- matrix(rnorm(300), nrow = 100)
kinship <- diag(100)
calc_covs(pheno, kinship)
```

```{r}
library(limmbo2)
limmbo2(kinship = kinship, pheno = pheno, seed = 1)
```

## 5-dimensional phenotype comparison of `calc_covs()` and `limmbo2()`

```{r}
pheno <- matrix(rnorm(500), nrow = 100)
system.time(calc_covs(pheno, kinship))
```

```{r}
system.time(limmbo2(kinship = kinship, pheno = pheno, seed = 1))
```

## Use nontrivial kinship with trivariate phenotype


```{r}
n <- 500
A <- matrix(runif(n ^ 2), ncol = n) 
kinship <- round(t(A) %*% A, 6)
pheno <- matrix(rnorm(1500), nrow = 500)

system.time(calc_covs(pheno, 
                      kinship, 
                      max_iter = 1000,
                      max_prec = 1 / 10000) -> cc_out)
cc_out
```

```{r}
system.time(limmbo2(kinship = kinship, 
                    pheno = pheno, 
                    S = 3, 
                    seed = 1) -> li_out)
li_out
```


