---
title: "Testing close linkage v. pleiotropy"
author: "Frederick Boehm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setting for close linkage v pleiotropy likelihood ratio test

Our setting involves a pair of traits, $Y_1$ and $Y_2$, each of which individually (univariately) associates with a single genomic region. The exact definition of a genomic region is imprecise; in practice, it may be as large as 4 or 5 Mb. We seek to distinguish whether $Y_1$ and $Y_2$ associations (in the genomic region) arise due to a single QTL or whether there are two two distinct loci, each of which associates with exactly one of the two traits. We recognize that more complicated association patterns are possible, but we neglect them in this test.










## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
